Synthesis of Statically Analyzable Accelerator Networks From Sequential Programs

### 2016-04-24 paper.tex
before merging related work and preliminaries

The rest of the paper is organized as follows:
Section~\ref{pre} provides preliminaries in process network
%intermediate representation of the sequential program,
and accelerator generation using HLS. Section~\ref{mainConvert} describes the main steps in transforming sequential programs
to networks of processes. When implemented on FPGAs, these networks may experience deadlock issues, which can be analyzed and resolved with the method developed in section~\ref{deadlock}. To make our general approach more concrete, we presents a specific algorithm which fits into our PN generation framework in section~\ref{exampleAlgo}.
This algorithm and its complementing optimizations are developed with FPGA as the target platform, thus the generated networks boast great performance when mapped onto the intended 
computing substrate (section~\ref{expEval}). Finally, related work is presented in section~\ref{relatedwork}, before we conclude the paper with discussion and future directions (section~\ref{conclude}).

@@

%This is illustrated in figure~\ref{fig:example}.
\begin{figure}[htp]
\begin{center}
\includegraphics[width=1.0\linewidth]{fig/dlmo.pdf}
\caption{Example PN for Deadlock Analysis
\label{fig:example}}
\end{center}
\vspace{-1.5em}
\end{figure}

@@

\begin{figure}[htp]
\begin{center}
\includegraphics[width=1.05\linewidth]{fig/burstMem.pdf}
\caption{Precedence Graph for Process Network
\label{fig:memBurst}}
\end{center}
\vspace{-2.0em}
\end{figure}

Figure~\ref{fig:memBurst} illustrates how multiple burst mode memory accesses can
create a deadlock. The reaction of the memory is modelled by a pop action
responding to the incoming request, and a push action producing one response data token. Meanwhile, the sending of memory requests and receiving of responses are decoupled and assigned to different slots in the process schedule. This is often done in HLS because of the long latency of memory accesses.
For burst mode requests associated with a loop, the request sending is moved to outside of the loop as shown in the figure. The symbolic variable $B$ in the weight of the $DepM$ edge signifies how many tokens have to be consumed/buffered before the response for the next request ($data2$) become accessible for the response receiver. Assume the arbiter for the memory subsystem implements a policy which can potentially take in and serve multiple ($K$) consecutive requests from one stream, $B$ can be as big as $BurstSize_{max}*K$. 
In the case of a round-robin arbiter, $B = BurstSize_{max}$ and for our experiment platform which uses AXI bus protocol, this number is 256. With the precedence graph, we can easily compute that the buffering for the two memory response channels needs to be increased to $B$ for the deadlocks to be resolved.

@@

%\vspace{-0.7em}
\begin{table}[htbp]
\caption{Benchmark Descriptions}
%\vspace{-1.2em}
\scriptsize
\centering
\begin{tabular}{| c | c | c | }
  \hline

@@

\hline
\multirow{2}{*}{}Floyd-& {Graph Algo. with Dynamic Programming}  & \multirow{2}{*}{$\approx$ 8 MB}  \\
%\cline{2-3}                                                                                                                                                    
 Warshall & Memory Accesses Depend on Computation       &   \\
%\cline{2-3}                                                                          
\hline                                                                                                           
\multirow{2}{*}{Knapsack}& Optimization with Dynamic Programming &\multirow{2}{*}{$\approx$ 5 MB}  \\
%\cline{2-3}                                                                                                                                                    
  &Memory Accesses Depend on Computation       &   \\
%\cline{2-3}                                                             
\hline
\multirow{2}{*}{}Depth-First&  Number of Nodes = 4000 & \multirow{2}{*}{$\approx$ 3 MB}  \\
%\cline{2-3}                                                                                                                                                    
 Search & Number of Neighbors per Node = 200       &   \\
%\cline{2-3}                                                                        
\hline            
\multirow{2}{*}{}SpMV& Matrix in Compressed Row Storage & \multirow{2}{*}{ $\approx$ 16 MB}  \\
%\cline{2-3}                                                                                                                                                    
 Multiply &Indirect Memory Addressing      &  \\
%\cline{2-3}

@@

The FPGA device used for our evaluation is the Zynq-7000 XC7Z020 FPGA
SoC from Xilinx, installed on the ZedBoard evaluation platform. The SoC contains two parts: an ARM-processor based processing
system (PS), and the programmable logic (PL). 
The baseline for our evaluation is the performance of each software kernel running
on the ARM core in the PS. 
The core is an out-of-order, dual-issue hard
processor running at 667MHz. The Zynq platform also provides two
options for the accelerators in PL to access the main memory subsystem: 
through the accelerator coherence port (ACP), or the high performance (HP) port. 
The former connects to the snoop control unit in the processing system and 
thus uses/modifies the processing system's on chip cache. The HP port connects
directly to the memory controller, thus often imposes a longer round-trip latency
for data requests.
\begin{comment}
which necessitates the flushing of cache lines 
by the processor if cached memory is accessed by the accelerator.
In either case, if memory states are also buffered in the PL with caches, they 
need to be explicitly pushed to the processing system side after the accelerator finishes running. 
As both the ACP
and the HP are slave ports, they provide no mechanisms to extract data from the programmable logic
when the ARM processor is running. The interaction between the generated accelerators
and the main pieces of the FPGA SoC is shown in figure~\ref{fig:ippf}.
\end{comment}

@@

\begin{figure*}[htp]
\begin{center}
\includegraphics[width=0.95\linewidth]{fig/perf.pdf}
\caption{Performance of Conventional Accelerators and Accelerator Networks, normalized to the hard ARM core. Implementations with caches use
Xilinx System Cache IP, configured to be 64KB and 2-way associative   
\label{fig:acpraw}}
\end{center}
\vspace{-2.7em}
\end{figure*}

In our study, Vivado HLS, a state-of-the-art high level synthesis tool provided by Xilinx,
is used to generate the accelerators for the original sequential code, as well as the processes created by our flow. With the target clock period set to 
8ns during HLS, the tightest timing constraints post place \& route implementations managed to meet range from 111 to 150MHz. 
All design points shown in this section 
use the highest achievable frequency as the actual operating clock frequency. 
%\vspace{-1.0em}
\subsection{Performance Comparisons}
In figure~\ref{fig:acpraw}, 
performance of the different implementations are presented. 
Conventional accelerators and dataflow accelerators with different memory subsystem
configurations are compared. All the numbers are normalized to the baseline.

In all four benchmarks, conventional accelerators directly generated from software kernels using the HLS flow
actually have lower performance than the hard processor. Even with
on-PL caches, 
these accelerators only manage to achieve throughput less than 50\% that of the baseline. 
The superscalar, 
out-of-order ARM core is capable of exploiting instruction level parallelism to a good extent 
and also has a high performance on-chip cache.
The additional parallelism extracted by the HLS tool is evidently not enough
to compensate for the clock frequency advantage the hard processor core has over the programmable logic and the longer data access
latency from the reconfigurable array.

With our methodology, the accelerator networks generated 
are rather competitive against the hard processor, even
without a reconfigurable cache. For SpMV multiply, knapsack and Floyd-Warshall, when the networks are directly connected to the memory through the ACP, 
the average performance is 2.3x that of the baseline---representing
an 8.4x gain over the conventional accelerators.
Upon the addition of caches,  
the average runtime of the accelerator networks was reduced by 18.7\%,  while that of the conventional
accelerators was cut by 45.4\%. The gap between their performance is thereby reduced from 8.4 to 5.6 times. This difference in improvement is due to conventional accelerators' sensitivity to the latency of data accesses, which is also manifested by its performance degradation of 40\% when the uncached HP port is used instead of ACP.

@@

Overall, for kernels suitable for FPGA acceleration, there is a significant performance advantage in using
an intermediate dataflow architectural template. If we compare the best results achieved with the
dataflow accelerators to that of the conventional accelerators, we see improvement
of 3.3 to 9.1 times, with an average of 5.6. 
%This performance advantage translates to energy savings as well, as illustrated in the next section.

%\subsection{Area and Energy Comparison}
%\vspace{-1.0em}
\subsection{Area comparison}
To quantify the impact of our proposed methodology on area, 
we have compared the FPGA resouce usage of conventional accelerators and
the dataflow accelerators.
Table~\ref{tab:areacom} shows the results, 
where each acclerator is complemented with two different memory subsystem configurations.
\renewcommand{\tabcolsep}{0.5pt}
\begin{table}[htbp]
%\vspace{-1.2em}
\caption{Resource Usage of Accelerators.}
%\vspace{-1.2em}
\scriptsize
\centering
\begin{tabular}{| c | c | c | c | c | c | c | c| }
  \hline            
  \multirow{2}{*}{} &  &   \multicolumn{3}{c|}{\bf ACP  } & \multicolumn{3}{c|}{\bf ACP + 64KB Cache }   \\
 \cline{3-8} 
 {\bf Benchmark}   &    & LUT& FFs& BRAM & LUT &   FFs      & BRAM    \\
  %
  %                 & LUT &FFs&  BRAM &  LUT &   FFs      & BRAM   \\
  \hline            
  \hline            
\multirow{3}{*}{}&Con.Acc  & 9873 &9116 &10 & 7918 & 6792 & 21  \\
\cline{2-8}                                                                                                                                                    
SpMV &Dataflow Acc       & 8577 & 8837& 10 & 6718  &6788 & 21\\
\cline{2-8}                                                                                                             
    Multiply   &\% change & -13.1 &-3.1 & 0 & -15.2  & -0.1 & 0  \\
  \hline                                                                                                           
\multirow{3}{*}{}&Con.Acc  & 7672 &7490 &8 & 6573 & 5885 & 21  \\
\cline{2-8}                                                                                                                                                    
Knapsack &Dataflow Acc       & 8089 & 8787& 8 & 6970  &7256 & 21\\
\cline{2-8}                                                                                                             
       &\% change & +5.4 &+17.3 & 0 & +6.0  & +23.3 & 0  \\
  \hline                                                                                                           
\multirow{3}{*}{}&Con.Acc  & 2491 &3528 &0 &3806  &4629  &19   \\
\cline{2-8}                                                                                                                                                    
Floyd- &Dataflow Acc       & 7659 &7210 &0  &8995  &8309 &19 \\
\cline{2-8}                                                                                                             
 Warshall      &\% change &+207.5  &+104.3 & 0 & +104.4  & +79.5  & 0  \\
  \hline                                                                                                           
\multirow{3}{*}{}&Con.Acc  & 4810 &4929 &4 & 4931 & 4594 & 21  \\
\cline{2-8}                                                                                                                                                    
DFS &Dataflow Acc       & 8509 & 7813& 4 & 7436  &6298 & 21\\
\cline{2-8}                                                                                                             
       &\% change & +76.9 &+58.5 & 0 & +50.8  & +37.1 & 0  \\
  \hline

@@

\section{Conclusions}
\label{conclude}
This paragraph will end the body of this sample document.
Remember that you might still have Acknowledgments or
Appendices; brief samples of these
follow.  There is still the Bibliography to deal with; and
we will make a disclaimer about that here: with the exception
of the reference to the \LaTeX\ book, the citations in
this paper are to articles which have nothing to
do with the present subject and are used as
examples only.
%\end{document}  % This is where a 'short' article might terminate

>>>>

The rest of the paper is organized as follows:
Section~\ref{pre} provides preliminaries in process network
%intermediate representation of the sequential program,
and accelerator generation using HLS. Section~\ref{mainConvert} describes the main steps in transforming sequential programs
to networks of processes. When implemented on FPGAs, these networks may experience deadlock issues, which can be analyzed and resolved with the method developed in section~\ref{deadlock}. To make our general approach more concrete, we presents a specific algorithm which fits into our PN generation framework in section~\ref{exampleAlgo}.
This algorithm and its complementing optimizations are developed with FPGA as the target platform, thus the generated networks boast great performance when mapped onto the intended 
computing substrate (section~\ref{expEval}). Finally, related work is presented in section~\ref{relatedwork}, before we conclude the paper (section~\ref{conclude}).
%with discussion and future directions (section~\ref{conclude}).

@@

%This is illustrated in figure~\ref{fig:example}.
\begin{figure}[htp]
\begin{center}
\includegraphics[width=1.0\linewidth]{fig/dlmo.pdf}
\caption{Example PN for Deadlock Analysis
\label{fig:example}}
\end{center}
\vspace{-2.0em}
\end{figure}

@@

Figure~\ref{fig:memBurst} illustrates how multiple burst mode memory accesses can
create a deadlock. The reaction of the memory is modelled by a pop action
responding to the incoming request, and a push action producing one response data token. Meanwhile, the sending of memory requests and receiving of responses are decoupled and assigned to different slots in the process schedule. This is often done in HLS because of the long latency of memory accesses.
For burst mode requests associated with a loop, the request sending is moved to outside of the loop as shown in the figure. The symbolic variable $B$ in the weight of the $DepM$ edge signifies how many tokens have to be consumed/buffered before the response for the next request ($data2$) become accessible for the response receiver. Assume the arbiter for the memory subsystem implements a policy which can potentially take in and serve multiple ($K$) consecutive requests from one stream, $B$ can be as big as $BurstSize_{max}*K$. 
In the case of a round-robin arbiter, $B = BurstSize_{max}$ and for our experiment platform which uses AXI bus protocol, this number is 256. With the precedence graph, we can easily compute that the buffering for the two memory response channels needs to be increased to $B$ for the deadlocks to be resolved.
\begin{figure}[htp]
\begin{center}
\includegraphics[width=1.05\linewidth]{fig/burstMem.pdf}
\caption{Deadlock in Burst Memory Accesses
\label{fig:memBurst}}
\end{center}
\vspace{-2.0em}
\end{figure}

@@

%\vspace{-0.7em}

The FPGA device used for our evaluation is Xilinx's Zynq-7000 XC7Z020 FPGA
SoC. It contains two parts: an ARM based processing
system (PS), and the programmable logic (PL). 
The baseline for our evaluation is the performance of each kernel executing
on the ARM core, which is an out-of-order, dual-issue hard processor running at 667MHz. Zynq also provides two
options for the accelerators in PL to access the main memory: 
through the accelerator coherence port (ACP), or the high performance (HP) port. 
ACP connects to the snoop control unit in the PS, and 
thus accesses the on-chip cache of the PS. On the other hand, the HP port connects
directly to the memory controller, thus often incurs longer latency
for data requests.
\begin{comment}
which necessitates the flushing of cache lines 
by the processor if cached memory is accessed by the accelerator.
In either case, if memory states are also buffered in the PL with caches, they 
need to be explicitly pushed to the processing system side after the accelerator finishes running. 
As both the ACP
and the HP are slave ports, they provide no mechanisms to extract data from the programmable logic
when the ARM processor is running. The interaction between the generated accelerators
and the main pieces of the FPGA SoC is shown in figure~\ref{fig:ippf}.
\end{comment}

\begin{table}[htbp]
\caption{Benchmark Descriptions}
%\vspace{-1.2em}
\scriptsize
\centering
\begin{tabular}{| c | c | c | }
  \hline

@@

\hline
\multirow{2}{*}{}Floyd-& {Graph Algo. with Dynamic Programming}  & \multirow{2}{*}{$\approx$ 8 MB}  \\
%\cline{2-3}                                                                                                                                                    
 Warshall & Memory Accesses Depend on Computation       &   \\
%\cline{2-3}                                                                          
\hline                                                                                                           
\multirow{2}{*}{Knapsack}& Optimization with Dynamic Programming &\multirow{2}{*}{$\approx$ 5 MB}  \\
%\cline{2-3}                                                                                                                                                    
  &Memory Accesses Depend on Computation       &   \\
%\cline{2-3}                                                             
\hline
\multirow{2}{*}{} &    & \multirow{2}{*}{$\approx$ 3 MB}  \\
%\cline{2-3}                                                                                                                                                    
   &         &   \\
%\cline{2-3}                                                                        
\hline            
\multirow{2}{*}{}SpMV& Matrix in Compressed Row Storage & \multirow{2}{*}{ $\approx$ 16 MB}  \\
%\cline{2-3}                                                                                                                                                    
 Multiply &Indirect Memory Addressing      &  \\
%\cline{2-3}

@@

@@

\begin{figure*}[htp]
\begin{center}
\includegraphics[width=0.95\linewidth]{fig/perfNew.pdf}
\caption{Performance of Conventional Accelerators and Accelerator Networks, normalized to the hard ARM core. Implementations with caches use
Xilinx System Cache IP, configured to be 64KB and 2-way associative   
\label{fig:acpraw}}
\end{center}
\vspace{-2.7em}
\end{figure*}

In our study, Vivado HLS, a state-of-the-art high level synthesis tool provided by Xilinx,
is used to create the accelerators for the original sequential code, as well as the processes generated by our flow. With the target clock period set to 
8ns during HLS, the tightest timing constraints post place \& route implementations managed to meet range from 111 to 150MHz. 
All design points shown in this section 
use the highest achievable frequency as the actual operating clock frequency. 
%\vspace{-1.0em}
\subsection{Performance Comparisons}
In figure~\ref{fig:acpraw}, 
performance of the different implementations are presented. 
Conventional accelerators and accelerator networks with different memory subsystem
configurations are compared. All the numbers are normalized to the baseline.

In all four benchmarks, conventional accelerators directly generated from software kernels actually have lower performance than the hard processor. Even with
on-PL caches, they can only achieve throughput less than 50\% that of the baseline.
The superscalar, out-of-order ARM core is capable of exploiting instruction level parallelism to a good extent and also has a high performance on-chip cache.
The additional parallelism exploited by the HLS flow is evidently not enough
to compensate for the clock frequency advantage the hard processor core has over the programmable logic and the longer data access latency from the reconfigurable array.

With our methodology, the accelerator networks generated 
are rather competitive against the hard processor, even
without a reconfigurable cache. When the networks are directly connected to the memory through the ACP, the average performance is 2.3x that of the baseline---an 8.4x gain over the conventional accelerators. 
With the addition of caches,  
the average runtime of the accelerator networks was reduced by 18.7\%,  while that of the conventional accelerators was cut by 45.4\%. The gap between their performance is thereby reduced from 8.4x to 5.6x. This difference in improvement is due to conventional accelerators' sensitivity to the latency of data accesses, which is also manifested by its performance degradation of 40\% when the uncached HP port is used instead of ACP.

@@

Overall, there is a significant performance advantage in the accelerator network as our algorithm effectively addresses the weakness of conventional accelerators. If we compare the best results achieved with the
accelerator networks to that of the conventional accelerators, we see improvement
of 3.3 to 9.1 times, with an average of 5.6. 
%This performance advantage translates to energy savings as well, as illustrated in the next section.

%\subsection{Area and Energy Comparison}
%\vspace{-1.0em}
\subsection{Area comparison}
To quantify the impact of our proposed methodology on area, 
we have compared the FPGA resouce usage of conventional accelerators and
the accelerator networks.
Table~\ref{tab:areacom} shows the results, 
where each acclerator is complemented with two different memory subsystem configurations.
\renewcommand{\tabcolsep}{0.5pt}
\begin{table}[htbp]
%\vspace{-1.2em}
\caption{Resource Usage of Accelerators.}
%\vspace{-1.2em}
\scriptsize
\centering
\begin{tabular}{| c | c | c | c | c | c | c | c| }
  \hline            
  \multirow{2}{*}{} &  &   \multicolumn{3}{c|}{\bf ACP  } & \multicolumn{3}{c|}{\bf ACP + 64KB Cache }   \\
 \cline{3-8} 
 {\bf Benchmark}   &    & LUT& FFs& BRAM & LUT &   FFs      & BRAM    \\
  %
  %                 & LUT &FFs&  BRAM &  LUT &   FFs      & BRAM   \\
  \hline            
  \hline

\multirow{3}{*}{}&Con.Acc  & 2491 &3528 &0 &3806  &4629  &19   \\
\cline{2-8}                                                                                                                                                    
Floyd- &Acc. Network       & 7659 &7210 &0  &8995  &8309 &19 \\
\cline{2-8}                                                                                                             
 Warshall      &\% change &+207.5  &+104.3 & 0 & +104.4  & +79.5  & 0  \\
  \hline

  \multirow{3}{*}{}&Con.Acc  & 7672 &7490 &8 & 6573 & 5885 & 21  \\
\cline{2-8}                                                                                                                                                    
Knapsack &Acc. Network       & 8089 & 8787& 8 & 6970  &7256 & 21\\
\cline{2-8}                                                                                                             
       &\% change & +5.4 &+17.3 & 0 & +6.0  & +23.3 & 0  \\
  \hline

\multirow{3}{*}{}&Con.Acc  & 4810 &4929 &4 & 4931 & 4594 & 21  \\
\cline{2-8}                                                                                                                                                    
DFS &Acc. Network       & 8509 & 7813& 4 & 7436  &6298 & 21\\
\cline{2-8}                                                                                                             
       &\% change & +76.9 &+58.5 & 0 & +50.8  & +37.1 & 0  \\
  \hline                                                                        
  \multirow{3}{*}{}&Con.Acc  & 9873 &9116 &10 & 7918 & 6792 & 21  \\
\cline{2-8}                                                                                                                                                    
SpMV &Acc. Network       & 8577 & 8837& 10 & 6718  &6788 & 21\\
\cline{2-8}                                                                                                             
    Multiply   &\% change & -13.1 &-3.1 & 0 & -15.2  & -0.1 & 0  \\
  \hline

@@

\section{Conclusions}
\label{conclude}
In this paper, we presented a framework with which sequential programs can be
transformed to networks of processes. Using HLS as our backend, they can then be implemented as accelerator networks on FPGA. Compare to accelerators directly synthesized from the original program, and software running on a hard CPU core with much higher clock frequency, our final implementations can achieve significantly better performance.
%\end{document}  % This is where a 'short' article might terminate
